---
title: "Prisons explore"
output: html_document
---

## First step to load packages and file locations etc.
```{r include=FALSE}
rm(list=ls())
project.folder = paste0(print(here::here()),'/')
source(paste0(project.folder,'create_folder_structure.R'))
source(paste0(functions.folder,'script_initiate.R'))
```

## Load specific plotting packages
```{r, message=FALSE}
# load required packages
library(data.table)
library(maptools)
library(mapproj)
library(rgeos)
library(rgdal)
library(RColorBrewer)
library(ggplot2)
library(raster)
library(sp)
library(plyr)
library(graticule)
library(zoo)
library(purrr)
library(cowplot)
```

## Prepare map structure
```{r}
# for map theme to plot in ggplot
theme_map = function(base_size=8 , base_family=""){
    require(grid)
    theme_bw(base_size=base_size,base_family=base_family) %+replace%
    theme(axis.line=element_blank(),
    axis.text=element_blank(),
    axis.ticks=element_blank(),
    axis.title=element_blank(),
    panel.background=element_blank(),
    panel.border=element_blank(),
    panel.grid=element_blank(),
    panel.margin=unit(0,"lines"),
    plot.background=element_blank(),
    legend.position = 'bottom'
    )
}
```

## Load US prison shapefile for mapping
```{r}
# load shapefile of entire United States by prison
us.main = readOGR(dsn=paste0(project.folder,"data/shapefiles/Prison_Boundaries/"),
                           layer="Prison_Boundaries_Edited")

us.main = spTransform(us.main, CRS("+proj=laea +lat_0=45 +lon_0=-100 +x_0=0 +y_0=0 +a=6370997 +b=6370997 +units=m +no_defs"))

# fortify to prepare for plotting in ggplot
map = fortify(us.main)

# extract data from shapefile
us.main@data$id = rownames(us.main@data) 
shapefile.data = us.main@data

# fix the weird Alabama mistake
shapefile.data = shapefile.data %>%
  dplyr::mutate(STATEFP=case_when(STATEFP=='10' & STATE=='AL'~ '01',
                               TRUE ~ STATEFP))

# merge selected data to map_create dataframe for colouring of ggplot
USA.df = merge(map, shapefile.data, by='id')
USA.df$prison_id = as.integer(as.character(USA.df$FID))
```

## Only choose prisons which are operational and with more than zero population
```{r}
shapefile.data = shapefile.data %>% 
  dplyr::filter(STATUS=='OPEN') %>% 
  dplyr::filter(POPULATION > 0) %>%
  dplyr::rename(Type=TYPE) %>%
  dplyr::mutate(Type1=case_when(Type=='COUNTY' ~ 'County',
                               Type=='LOCAL' ~ 'Local',
                               Type=='FEDERAL' ~ 'Federal',
                               Type=='STATE' ~ 'State',
                               TRUE ~ 'Other'))
```

## How many prisons included in total?
```{r}
shapefile.data %>% tally()
```

# How many prisons included by type?
```{r}
shapefile.data %>% group_by(Type) %>% tally()
```

# How many prisons by state?
```{r}
shapefile.data %>% group_by(STATE) %>% tally() %>% arrange(desc(n))
```

# How many incarcerated people?
```{r}
shapefile.data %>% tally(POPULATION)
```

# How many incarcerated people included by type?
```{r}
shapefile.data %>% group_by(Type)  %>% tally(POPULATION)
```

# How many incarcerated people by state?
```{r}
shapefile.data %>% group_by(STATE) %>% tally(POPULATION) %>% arrange(desc(n))
```