---
title: "Figure 1"
output: html_document
---

## First step to load packages and file locations etc.
```{r include=FALSE}
rm(list=ls())
project.folder = paste0(print(here::here()),'/')
source(paste0(project.folder,'create_folder_structure.R'))
source(paste0(functions.folder,'script_initiate.R'))
```

## Load specific plotting packages
```{r, message=FALSE}
# load required packages
library(data.table)
library(maptools)
library(mapproj)
library(rgeos)
library(rgdal)
library(RColorBrewer)
library(ggplot2)
library(raster)
library(sp)
library(plyr)
library(graticule)
library(zoo)
library(purrr)
library(cowplot)
```

## Load WBGT regression file
```{r}
dat.wbgt.summarised.regression = readRDS(paste0(wbgt.folder,'weighted_area_raster_prison_wbgtmax_daily_',start_year_wbgt,'_',end_year_wbgt,'_regression_analysis.rds'))
```

## Prepare map structure
```{r}
# for map theme to plot in ggplot
theme_map = function(base_size=10, base_family=""){
    require(grid)
    theme_bw(base_size=base_size,base_family=base_family) %+replace%
    theme(axis.line=element_blank(),
    axis.text=element_blank(),
    axis.ticks=element_blank(),
    axis.title=element_blank(),
    panel.background=element_blank(),
    panel.border=element_blank(),
    panel.grid=element_blank(),
    panel.margin=unit(0,"lines"),
    plot.background=element_blank(),
    legend.position = 'bottom'
    )
}
```

## Load US prison shapefile to load details and mapping
```{r}
# load shapefile of entire United States by prison
us.main = readOGR(dsn=paste0(project.folder,"data/shapefiles/Prison_Boundaries/"),
                           layer="Prison_Boundaries_Edited")

us.main = spTransform(us.main, CRS("+proj=laea +lat_0=45 +lon_0=-100 +x_0=0 +y_0=0 +a=6370997 +b=6370997 +units=m +no_defs"))

# fortify to prepare for plotting in ggplot
map = fortify(us.main)

# extract data from shapefile
us.main@data$id = rownames(us.main@data) 
shapefile.data = us.main@data

# fix the weird Alabama mistake
shapefile.data = shapefile.data %>%
  dplyr::mutate(STATEFP=case_when(STATEFP=='10' & STATE=='AL'~ '01',
                               TRUE ~ STATEFP))

# merge selected data to map_create dataframe for colouring of ggplot
USA.df = merge(map, shapefile.data, by='id')
USA.df$prison_id = as.integer(as.character(USA.df$FID))
```

## Load WBGT over time file
```{r}
dat.wbgt.summarised = readRDS(paste0(wbgt.folder,'weighted_area_raster_prison_wbgtmax_daily_',start_year_wbgt,'_',end_year_wbgt,'_over_time.rds'))
```

## Merge prison over time file with shapefile data then summarise by state and year weighted by prison population
```{r}
dat.wbgt.summarised.merged.weighted.prison = left_join(dat.wbgt.summarised,shapefile.data, by=c('prison_id'='FID')) %>% 
  dplyr::filter(STATUS=='OPEN') %>% 
  dplyr::filter(POPULATION > 0) %>%
  dplyr::group_by(STATE,STATEFP,TYPE,year) %>%
  dplyr::summarise(wbgt_26 = weighted.mean(wbgt_26,POPULATION),
            wbgt_28 = weighted.mean(wbgt_28,POPULATION),
            wbgt_30 = weighted.mean(wbgt_30,POPULATION),
            wbgt_35 = weighted.mean(wbgt_35,POPULATION),
            POPULATION = sum(POPULATION)) %>%
  dplyr::mutate(STATE=factor(STATE, levels=rev(sort(unique(STATE))))) %>%
  dplyr::mutate(wbgt_26_pop = wbgt_26 * POPULATION,
                wbgt_28_pop = wbgt_28 * POPULATION,
                wbgt_30_pop = wbgt_30 * POPULATION,
                wbgt_35_pop = wbgt_35 * POPULATION)
```

## Take 5-year average of most recent 5 years
```{r}
dat.wbgt.summarised.merged.weighted.prison.5year.average = dat.wbgt.summarised.merged.weighted.prison %>%
  dplyr::filter(year%in%years_wbgtmax[c((length(years_wbgtmax)-4):length(years_wbgtmax))]) %>%
  dplyr::group_by(STATE,STATEFP,TYPE) %>%
  dplyr::summarise(wbgt_26_pop_mean = mean(wbgt_26_pop),
                   wbgt_28_pop_mean = mean(wbgt_28_pop),
                   wbgt_30_pop_mean = mean(wbgt_30_pop),
                   wbgt_35_pop_mean = mean(wbgt_35_pop)) %>%
  dplyr::rename(Type=TYPE) %>%
  dplyr::mutate(Type=case_when(Type=='COUNTY' ~ 'County',
                               Type=='LOCAL' ~ 'Local',
                               Type=='FEDERAL' ~ 'Federal',
                               Type=='STATE' ~ 'State',
                               TRUE ~ 'Other')) %>%
  dplyr::mutate(Type=as.factor(Type)) %>%
  dplyr::mutate(Type=factor(Type,levels=c('Federal','State','County','Local','Other')))
```

## Plot barchart by state stacked by prison type
```{r}
library(MetBrewer)

plot_bar_chart = function(threshold_chosen,legend_chosen=1){ 
  
  # calculate rank of bars
  dat.rank = dat.wbgt.summarised.merged.weighted.prison.5year.average %>%
    group_by(STATE) %>%
    dplyr::summarise(sum=sum(get(paste0('wbgt_',threshold_chosen,'_pop_mean')))) %>%
    dplyr::arrange(sum) %>%
    dplyr::mutate(rank=row_number())
  
  dat.plot = left_join(dat.wbgt.summarised.merged.weighted.prison.5year.average,dat.rank)

 p = ggplot() + 
   geom_bar(data=dat.plot, 
                aes(x=reorder(STATE, rank),
                              y=get(paste0('wbgt_',threshold_chosen,'_pop_mean')),fill=Type),stat = "identity") +
   ylab(paste0('Mean annual number of dangerous hot-humid person-days\nfor incarcerated people during ',(end_year_wbgt-4),'-',end_year_wbgt)) + 
   xlab('State') + 
   scale_y_continuous(label=scales::comma) + 
   coord_flip() +
   scale_fill_manual(values=met.brewer('Lakota')) + 
   theme_bw() + theme(text = element_text(size = 10), 
   panel.grid.major = element_blank(),axis.text.x = element_text(angle=0 , size=8, vjust=0.5), 
   plot.title = element_text(hjust = 0.5),panel.background = element_blank(), 
   panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"), 
   panel.border = element_rect(colour = "black"),strip.background = element_blank(), 
   legend.position = 'bottom',legend.justification='center', 
   legend.background = element_rect(fill="white", size=.5, linetype="dotted")) 

 return(p) 
}

plot.bar.chart.26 = plot_bar_chart(26,1)
plot.bar.chart.28 = plot_bar_chart(28,1)
plot.bar.chart.30 = plot_bar_chart(30,1)
```

## Load US state shapefile for mapping
```{r}
us = readOGR(dsn=paste0(data.folder,"shapefiles/states"),layer="states")

# reproject shapefile
us_aea = spTransform(us, CRS("+proj=laea +lat_0=45 +lon_0=-100 +x_0=0 +y_0=0 +a=6370997 +b=6370997 +units=m +no_defs"))
us_aea = us
us_aea@data$id = rownames(us_aea@data)

# only keep mainland US States
us_aea = us_aea[us_aea$STATE_FIPS %in% states_included,]
us_aea = spTransform(us_aea, CRS("+proj=laea +lat_0=45 +lon_0=-100 +x_0=0 +y_0=0 +a=6370997 +b=6370997 +units=m +no_defs"))

us_aea = fortify(us_aea)
```

## Prepare summary maps
```{r}
# merge map with mean values
USA.df.merged = left_join(USA.df,dat.wbgt.summarised.regression,by='prison_id')
USA.df.merged = with(USA.df.merged, USA.df.merged[order(id,order),])
```

## Prepare colors and legends of maps
```{r}
legend_create_count_start_year = function(threshold)(paste0('Number of dangerous hot-humid days\nin ',start_year_wbgt))
legend_create_count_end_year = function(threshold)(paste0('Number of dangerous hot-humid days\nin ',end_year_wbgt))
legend_create_count_last_5_years = function(threshold)(paste0('Mean annual number of\ndangerous hot-humid days during ',(end_year_wbgt-4),'-',end_year_wbgt))
legend_create_annual_change = function(threshold)(paste0('Annual change during\n',start_year_wbgt,'-',end_year_wbgt ))
legend_create_total_change = function(threshold)(paste0('Total change in annual number of dangerous hot-humid days\nduring ',start_year_wbgt,'-',end_year_wbgt))
```

## Only take open prisons and whose population we know
```{r}
USA.df.merged = USA.df.merged %>% 
  filter(STATUS=='OPEN') %>% filter(POPULATION > 0) 
```

## Plot count in last 5 years map 
```{r}
plot_last_5_years = function(threshold_chosen,legend_chosen=0){
  p = ggplot() +
    geom_polygon(data=us_aea,aes(x=long,y=lat,group=group),fill='grey35',color='black',size=0.5) +
    geom_polygon(data=USA.df.merged,aes(x=long,y=lat,group=group,
                                        color=get(paste0('wbgt_',threshold_chosen,'_',(end_year_wbgt-4),'_',end_year_wbgt))),size=1.2) +
    scale_color_gradientn(colors=rev(MetBrewer::met.brewer('Tam')), limits = c(0,max(USA.df.merged$wbgt_28_2016_2020))) +
    coord_fixed() + xlab('') + ylab('') + theme_bw() +
    theme(panel.grid.major = element_blank(),
        axis.text.x = element_text(angle=90), axis.ticks.x=element_blank(),legend.text=element_text(size=5),
        panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"),
        panel.border = element_rect(colour = "black"),strip.background = element_blank(),legend.justification='center',
        legend.position = 'bottom', legend.background = element_rect(fill="white", size=.5, linetype="dotted")) +
    theme_map()
  
  if(legend_chosen==0){
    p = p +
          guides(color=FALSE)
  }
  if(legend_chosen==1){
    p = p +
          guides(color = guide_colorbar(direction = "horizontal", title.position="left",barwidth = 10,barheight = 1,title.vjust = 0.8, 
                                  title = legend_create_count_last_5_years(threshold_chosen)),legend.text=element_text(size=5)) 
    }  

    return(p)

}

plot.count.end.year.legend = cowplot::get_legend(plot_last_5_years(26,1))

plot.count.last.5.years.26 = plot_last_5_years(26,1)
plot.count.last.5.years.28 = plot_last_5_years(28,1)
plot.count.last.5.years.30 = plot_last_5_years(30,1)
```

## 26°C: Plot
```{r}
pdf(paste0(wbgtmax.folder,'Figure_1_26.pdf'),paper='a4r',width=0,height=0)
plot_grid(plot.bar.chart.26, plot.count.last.5.years.26, 
          label_size = 14 ,
          label_x = -0.02, label_y = 1,
          labels = c('(a)', '(b)'))
dev.off()

plot_grid(plot.bar.chart.26, plot.count.last.5.years.26, 
          label_size = 14 ,
          label_x = -0.02, label_y = 1,
          labels = c('(a)', '(b)'))
```

## 28°C: Plot
```{r}
pdf(paste0(wbgtmax.folder,'Figure_1_28.pdf'),paper='a4r',width=0,height=0)
plot_grid(plot.bar.chart.28, plot.count.last.5.years.28, 
          label_size = 14 ,
          label_x = -0.02, label_y = 1,
          labels = c('(a)', '(b)'))
dev.off()

plot_grid(plot.bar.chart.28, plot.count.last.5.years.28, 
          label_size = 14 ,
          label_x = -0.02, label_y = 1,
          labels = c('(a)', '(b)'))
```

## 30°C: Plot
```{r}
pdf(paste0(wbgtmax.folder,'Figure_1_30.pdf'),paper='a4r',width=0,height=0)
plot_grid(plot.bar.chart.30, plot.count.last.5.years.30, 
          label_size = 14 ,
          label_x = -0.02, label_y = 1,
          labels = c('(a)', '(b)'))
dev.off()

plot_grid(plot.bar.chart.30, plot.count.last.5.years.30, 
          label_size = 14 ,
          label_x = -0.02, label_y = 1,
          labels = c('(a)', '(b)'))
```

## Save output data for Figure 1a
```{r}
dat.wbgt.summarised.merged.weighted.prison.5year.average.rounded = dat.wbgt.summarised.merged.weighted.prison.5year.average %>%
  mutate(wbgt_26_pop_mean=round(wbgt_26_pop_mean),
         wbgt_28_pop_mean=round(wbgt_28_pop_mean),
         wbgt_30_pop_mean=round(wbgt_30_pop_mean),
         wbgt_35_pop_mean=round(wbgt_35_pop_mean))

dat.wbgt.summarised.merged.weighted.prison.5year.average.rounded.total = dat.wbgt.summarised.merged.weighted.prison.5year.average.rounded %>%
  group_by(STATE,STATEFP) %>%
  dplyr::summarise(wbgt_26_pop_mean=sum(wbgt_26_pop_mean),
         wbgt_28_pop_mean=sum(wbgt_28_pop_mean),
         wbgt_30_pop_mean=sum(wbgt_30_pop_mean),
         wbgt_35_pop_mean=sum(wbgt_35_pop_mean)) %>%
  mutate(Type='Total') %>%
dplyr::relocate(Type)

dat.wbgt.summarised.merged.weighted.prison.5year.average.rounded = rbind(dat.wbgt.summarised.merged.weighted.prison.5year.average.rounded,
                                                                         dat.wbgt.summarised.merged.weighted.prison.5year.average.rounded.total) %>%
  dplyr::arrange(STATEFP) %>%
  dplyr::select(STATE,STATEFP,Type,wbgt_28_pop_mean)

readr::write_csv(dat.wbgt.summarised.merged.weighted.prison.5year.average.rounded ,paste0(wbgtmax.folder,'Figure_1a.csv'))
```

## Save output data for Figure 1b
```{r}
dat.wbgt.summarised.regression.output = dat.wbgt.summarised.regression %>%
  dplyr::select(prison_id,wbgt_28_2016_2020) %>%
  left_join(shapefile.data,by=c('prison_id'='FID')) %>% 
  filter(STATUS=='OPEN') %>% filter(POPULATION > 0) %>%
  dplyr::select(prison_id,wbgt_28_2016_2020,NAME,STATE,STATEFP,COUNTY,CITY,TYPE,POPULATION,CAPACITY,VAL_DATE)

readr::write_csv(dat.wbgt.summarised.regression.output ,paste0(wbgtmax.folder,'Figure_1b.csv'))
```