---
title: "Figure 1 additional values for press release"
output: html_document
---

## First step to load packages and file locations etc.
```{r include=FALSE}
rm(list=ls())
project.folder = paste0(print(here::here()),'/')
source(paste0(project.folder,'create_folder_structure.R'))
source(paste0(functions.folder,'script_initiate.R'))
```

## Load specific plotting packages
```{r, message=FALSE}
# load required packages
library(data.table)
library(maptools)
library(mapproj)
library(rgeos)
library(rgdal)
library(RColorBrewer)
library(ggplot2)
library(raster)
library(sp)
library(plyr)
library(graticule)
library(zoo)
library(purrr)
library(cowplot)
```

## Load WBGT regression file
```{r}
dat.wbgt.summarised.regression = readRDS(paste0(wbgt.folder,'weighted_area_raster_prison_wbgtmax_daily_',start_year_wbgt,'_',end_year_wbgt,'_regression_analysis.rds'))
```

## Prepare map structure
```{r}
# for map theme to plot in ggplot
theme_map = function(base_size=10, base_family=""){
    require(grid)
    theme_bw(base_size=base_size,base_family=base_family) %+replace%
    theme(axis.line=element_blank(),
    axis.text=element_blank(),
    axis.ticks=element_blank(),
    axis.title=element_blank(),
    panel.background=element_blank(),
    panel.border=element_blank(),
    panel.grid=element_blank(),
    panel.margin=unit(0,"lines"),
    plot.background=element_blank(),
    legend.position = 'bottom'
    )
}
```

## Load US prison shapefile to load details and mapping
```{r}
# load shapefile of entire United States by prison
us.main = readOGR(dsn=paste0(project.folder,"data/shapefiles/Prison_Boundaries/"),
                           layer="Prison_Boundaries_Edited")

us.main = spTransform(us.main, CRS("+proj=laea +lat_0=45 +lon_0=-100 +x_0=0 +y_0=0 +a=6370997 +b=6370997 +units=m +no_defs"))

# fortify to prepare for plotting in ggplot
map = fortify(us.main)

# extract data from shapefile
us.main@data$id = rownames(us.main@data) 
shapefile.data = us.main@data

# fix the weird Alabama mistake
shapefile.data = shapefile.data %>%
  dplyr::mutate(STATEFP=case_when(STATEFP=='10' & STATE=='AL'~ '01',
                               TRUE ~ STATEFP))

# merge selected data to map_create dataframe for colouring of ggplot
USA.df = merge(map, shapefile.data, by='id')
USA.df$prison_id = as.integer(as.character(USA.df$FID))
```

## Load WBGT over time file
```{r}
dat.wbgt.summarised = readRDS(paste0(wbgt.folder,'weighted_area_raster_prison_wbgtmax_daily_',start_year_wbgt,'_',end_year_wbgt,'_over_time.rds'))
```

# Last 5 years

## Only take for last 5 years and sum up number of days
```{r}
dat.wbgt.summarised.last.5.years = dat.wbgt.summarised %>%
  dplyr::filter(year%in%years_wbgtmax[c((length(years_wbgtmax)-4):length(years_wbgtmax))]) %>%
  # dplyr::mutate(wbgt_28 = ifelse(wbgt_28>1,1,0)) %>%
  dplyr::group_by(prison_id) %>%
  dplyr::summarise(wbgt_28 = sum(wbgt_28))
```

## Merge prison over time file with shapefile data 
```{r}
dat.wbgt.summarised.merged.weighted.prison = left_join(dat.wbgt.summarised.last.5.years,shapefile.data, by=c('prison_id'='FID')) %>% 
  dplyr::filter(STATUS=='OPEN') %>% 
  dplyr::filter(POPULATION > 0) 
```

## Number of incarcerated people who experience at least 1 day of WBGT > 28 during 2016-2020
```{r}
dat.wbgt.summarised.merged.weighted.prison %>%
  dplyr::filter(wbgt_28 >= 1) %>%
  dplyr::summarise(population = sum(POPULATION))
```

## Which carceral facilities experienced at least 1 day of WBGT > 28 during 2016-2020?
```{r}
prison.ids = dat.wbgt.summarised.merged.weighted.prison %>%
  dplyr::filter(wbgt_28 >= 1) %>%
  pull(prison_id) %>%
  unique() %>%
  sort()
```

## Average number of days WBGT > 28 in each year during 2016-2020 for incarcerated people who experience at least 1 day of WBGT > 28 during 2016-2020
```{r}
dat.wbgt.summarised.merged.weighted.prison %>%
  dplyr::filter(prison_id %in% prison.ids) %>%
  dplyr::summarise(wbgt_28 = mean(wbgt_28))
```

# First 5 years

## Only take for last 5 years and sum up number of days
```{r}
dat.wbgt.summarised.first.5.years = dat.wbgt.summarised %>%
  dplyr::filter(year%in%years_wbgtmax[c(1:5)]) %>%
  # dplyr::mutate(wbgt_28 = ifelse(wbgt_28>1,1,0)) %>%
  dplyr::group_by(prison_id) %>%
  dplyr::summarise(wbgt_28 = sum(wbgt_28))
```

## Merge prison over time file with shapefile data 
```{r}
dat.wbgt.summarised.merged.weighted.prison = left_join(dat.wbgt.summarised.first.5.years,shapefile.data, by=c('prison_id'='FID')) %>% 
  dplyr::filter(STATUS=='OPEN') %>% 
  dplyr::filter(POPULATION > 0) 
```

## Number of incarcerated people who experience at least 1 day of WBGT > 28 during 1982-1986
```{r}
dat.wbgt.summarised.merged.weighted.prison %>%
  dplyr::filter(wbgt_28 >= 1) %>%
  dplyr::summarise(population = sum(POPULATION))
```

## Which carceral facilities experienced at least 1 day of WBGT > 28 during 1982-1986?
```{r}
prison.ids = dat.wbgt.summarised.merged.weighted.prison %>%
  dplyr::filter(wbgt_28 >= 1) %>%
  pull(prison_id) %>%
  unique() %>%
  sort()
```

## Average number of days WBGT > 28 in each year during 1982-1986 for incarcerated people who experience at least 1 day of WBGT > 28 during 1982-1986
```{r}
dat.wbgt.summarised.merged.weighted.prison %>%
  dplyr::filter(prison_id %in% prison.ids) %>%
  dplyr::summarise(wbgt_28 = mean(wbgt_28))
```
