---
title: "Figure 2"
output: html_document
---

## First step to load packages and file locations etc.
```{r include=FALSE}
rm(list=ls())
project.folder = paste0(print(here::here()),'/')
source(paste0(project.folder,'create_folder_structure.R'))
source(paste0(functions.folder,'script_initiate.R'))
```

## Load specific plotting packages
```{r, message=FALSE}
# load required packages
library(data.table)
library(maptools)
library(mapproj)
library(rgeos)
library(rgdal)
library(RColorBrewer)
library(ggplot2)
library(raster)
library(sp)
library(plyr)
library(graticule)
library(zoo)
library(purrr)
library(cowplot)
```

## Load WBGT regression file
```{r}
dat.wbgt.summarised.regression = readRDS(paste0(wbgt.folder,'weighted_area_raster_prison_wbgtmax_daily_',start_year_wbgt,'_',end_year_wbgt,'_regression_analysis.rds'))
```

## Prepare map structure
```{r}
# for map theme to plot in ggplot
theme_map = function(base_size=9, base_family=""){
    require(grid)
    theme_bw(base_size=base_size,base_family=base_family) %+replace%
    theme(axis.line=element_blank(),
    axis.text=element_blank(),
    axis.ticks=element_blank(),
    axis.title=element_blank(),
    panel.background=element_blank(),
    panel.border=element_blank(),
    panel.grid=element_blank(),
    panel.margin=unit(0,"lines"),
    plot.background=element_blank(),
    legend.position = 'bottom'
    )
}
```

## Load US prison shapefile for mapping
```{r}
# load shapefile of entire United States by prison
us.main = readOGR(dsn=paste0(project.folder,"data/shapefiles/Prison_Boundaries/"),
                           layer="Prison_Boundaries_Edited")

us.main = spTransform(us.main, CRS("+proj=laea +lat_0=45 +lon_0=-100 +x_0=0 +y_0=0 +a=6370997 +b=6370997 +units=m +no_defs"))

# fortify to prepare for plotting in ggplot
map = fortify(us.main)

# extract data from shapefile
us.main@data$id = rownames(us.main@data) 
shapefile.data = us.main@data

# fix the weird Alabama mistake
shapefile.data = shapefile.data %>%
  dplyr::mutate(STATEFP=case_when(STATEFP=='10' & STATE=='AL'~ '01',
                               TRUE ~ STATEFP))

# merge selected data to map_create dataframe for colouring of ggplot
USA.df = merge(map, shapefile.data, by='id')
USA.df$prison_id = as.integer(as.character(USA.df$FID))
```

## Load WBGT over time file
```{r}
dat.wbgt.summarised = readRDS(paste0(wbgt.folder,'weighted_area_raster_prison_wbgtmax_daily_',start_year_wbgt,'_',end_year_wbgt,'_over_time.rds'))
```

## Merge prison over time file with shapefile data then summarise by state and year weighted by prison population
```{r}
dat.wbgt.summarised.merged.weighted.prison = left_join(dat.wbgt.summarised,shapefile.data, by=c('prison_id'='FID')) %>% 
  dplyr::filter(STATUS=='OPEN') %>% 
  dplyr::filter(POPULATION > 0) %>%
  dplyr::group_by(STATE,STATEFP,year) %>%
  dplyr::summarise(wbgt_26 = weighted.mean(wbgt_26,POPULATION),
            wbgt_28 = weighted.mean(wbgt_28,POPULATION),
            wbgt_30 = weighted.mean(wbgt_30,POPULATION),
            wbgt_35 = weighted.mean(wbgt_35,POPULATION)) 

STATES = dat.wbgt.summarised.merged.weighted.prison %>% pull(STATE) %>% unique()

dat.wbgt.summarised.merged.weighted.prison.national = left_join(dat.wbgt.summarised,shapefile.data, by=c('prison_id'='FID')) %>% 
  dplyr::filter(STATUS=='OPEN') %>% 
  dplyr::filter(POPULATION > 0) %>%
  dplyr::group_by(year) %>%
  dplyr::summarise(wbgt_26 = weighted.mean(wbgt_26,POPULATION),
            wbgt_28 = weighted.mean(wbgt_28,POPULATION),
            wbgt_30 = weighted.mean(wbgt_30,POPULATION),
            wbgt_35 = weighted.mean(wbgt_35,POPULATION)) %>%
  dplyr::mutate(STATE='USA',STATEFP='99') %>%
  dplyr::relocate(STATE) %>%
  dplyr::relocate(STATEFP)

dat.wbgt.summarised.merged.weighted.prison = 
  rbind(dat.wbgt.summarised.merged.weighted.prison,dat.wbgt.summarised.merged.weighted.prison.national) %>%
  mutate(STATE=factor(STATE, levels=rev(c('USA',STATES))))
```

## Load state file over time
```{r}
dat.wbgt.summarised.merged.weighted.state = readRDS(paste0(wbgt.folder,'weighted_area_raster_state_wbgtmax_daily_',start_year_wbgt,'_',end_year_wbgt,'_over_time.rds'))
```

## Merge prison vs. state to compare
```{r}
dat.wbgt.summarised.merged.weighted.prison = dat.wbgt.summarised.merged.weighted.prison %>%
  dplyr::rename(wbgt_26_prison=wbgt_26,wbgt_28_prison=wbgt_28,wbgt_30_prison=wbgt_30,wbgt_35_prison=wbgt_35) %>%
  dplyr::rename(state=STATEFP)

dat.wbgt.summarised.merged.weighted.state = dat.wbgt.summarised.merged.weighted.state %>%
  dplyr::rename(wbgt_26_state=wbgt_26,wbgt_28_state=wbgt_28,wbgt_30_state=wbgt_30,wbgt_35_state=wbgt_35)

dat.wbgt.summarised.merged.weighted.prison.state = left_join(dat.wbgt.summarised.merged.weighted.prison,
                                                             dat.wbgt.summarised.merged.weighted.state,
                                                             by=c('state','year')) %>%
  dplyr::mutate(wbgt_26_diff = wbgt_26_prison - wbgt_26_state,
                wbgt_28_diff = wbgt_28_prison - wbgt_28_state,
                wbgt_30_diff = wbgt_30_prison - wbgt_30_state,
                wbgt_35_diff = wbgt_35_prison - wbgt_35_state)
```

## Load US state shapefile for mapping
```{r}
us = readOGR(dsn=paste0(data.folder,"shapefiles/states"),layer="states")

# reproject shapefile
us_aea = spTransform(us, CRS("+proj=laea +lat_0=45 +lon_0=-100 +x_0=0 +y_0=0 +a=6370997 +b=6370997 +units=m +no_defs"))
us_aea = us
us_aea@data$id = rownames(us_aea@data)

# only keep mainland US States
us_aea = us_aea[us_aea$STATE_FIPS %in% states_included,]
us_aea = spTransform(us_aea, CRS("+proj=laea +lat_0=45 +lon_0=-100 +x_0=0 +y_0=0 +a=6370997 +b=6370997 +units=m +no_defs"))

us_aea = fortify(us_aea)
```

## Prepare summary maps
```{r}
# merge map with mean values
USA.df.merged = left_join(USA.df,dat.wbgt.summarised.regression,by='prison_id')
USA.df.merged = with(USA.df.merged, USA.df.merged[order(id,order),])
```

## Prepare colors and legends of maps
```{r}
legend_create_count_start_year = function(threshold)(paste0('Number of dangerous hot-humid days\nin ',start_year_wbgt))
legend_create_count_end_year = function(threshold)(paste0('Number of dangerous hot-humid days\nin ',end_year_wbgt))
legend_create_count_last_5_years = function(threshold)(paste0('Mean annual number of dangerous hot-humid days\nduring ',(end_year_wbgt-4),'-',end_year_wbgt))
legend_create_annual_change = function(threshold)(paste0('Annual change during\n',start_year_wbgt,'-',end_year_wbgt ))
legend_create_total_change = function(threshold)(paste0('Total change in annual number of\ndangerous hot-humid days during ',start_year_wbgt,'-',end_year_wbgt))
```

## Only take open prisons and whose population we know
```{r}
USA.df.merged = USA.df.merged %>% 
  filter(STATUS=='OPEN') %>% filter(POPULATION > 0) 
```

## Plot count in start year map 
```{r}
plot_start_year = function(threshold_chosen,legend_chosen=0){
  p = ggplot() +
    geom_polygon(data=us_aea,aes(x=long,y=lat,group=group),fill='grey35',color='black',size=0.5) +
    geom_polygon(data=USA.df.merged,aes(x=long,y=lat,group=group,color=get(paste0('wbgt_',threshold_chosen,'_',start_year_wbgt))),size=1) +
    scale_color_gradientn(colors=(MetBrewer::met.brewer('OKeeffe2')), limits = c(0,max(USA.df.merged$wbgt_28_1982))) +
    coord_fixed() + xlab('') + ylab('') + theme_bw() +
    ggtitle(paste0(threshold_chosen,'°C')) + 
    theme(panel.grid.major = element_blank(),
        axis.text.x = element_text(angle=90), axis.ticks.x=element_blank(),legend.text= element_text(size=5),
        panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"),
        panel.border = element_rect(colour = "black"),strip.background = element_blank(),legend.justification='center',
        plot.title = element_text(hjust = 0.5),
        legend.position = 'bottom', legend.background = element_rect(fill="white", size=.5, linetype="dotted")) +
    theme_map() + 
    theme(plot.title = element_text(hjust = 0.5))
  
  if(legend_chosen==0){
    p = p +
          guides(color=FALSE)
  }
  if(legend_chosen==1){
    p = p +
          guides(color = guide_colorbar(direction = "horizontal", title.position="left",barwidth = 10,barheight = 1,title.vjust = 0.8, 
                                  title = legend_create_count_start_year(threshold_chosen)),legend.text=element_text(size=5))  
    }

  return(p)
}

plot.count.start.year.legend = cowplot::get_legend(plot_start_year(26,1))

plot.count.start.year.26 = plot_start_year(26,0)
plot.count.start.year.28 = plot_start_year(28,0)
plot.count.start.year.30 = plot_start_year(30,0)
```

## Plot count in end year map 
```{r}
plot_end_year = function(threshold_chosen,legend_chosen=0){
  p = ggplot() +
    geom_polygon(data=us_aea,aes(x=long,y=lat,group=group),fill='grey35',color='black',size=0.5) +
    geom_polygon(data=USA.df.merged,aes(x=long,y=lat,group=group,color=get(paste0('wbgt_',threshold_chosen,'_',end_year_wbgt))),size=1) +
    scale_color_gradientn(colors=MetBrewer::met.brewer('OKeeffe2'), limits = c(0,max(USA.df.merged$wbgt_28_1982))) +
    coord_fixed() + xlab('') + ylab('') + theme_bw() +
    theme(panel.grid.major = element_blank(),
        axis.text.x = element_text(angle=90), axis.ticks.x=element_blank(),legend.text=element_text(size=5),
        panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"),
        panel.border = element_rect(colour = "black"),strip.background = element_blank(),legend.justification='center',
        legend.position = 'bottom', legend.background = element_rect(fill="white", size=.5, linetype="dotted")) +
    theme_map()
  
  if(legend_chosen==0){
    p = p +
          guides(color=FALSE)
  }
  if(legend_chosen==1){
    p = p +
          guides(color = guide_colorbar(direction = "horizontal", title.position="left",barwidth = 10,barheight = 1,title.vjust = 0.8, 
                                  title = legend_create_count_end_year(threshold_chosen)),legend.text=element_text(size=5)) 
    }  

    return(p)

}

plot.count.end.year.legend = cowplot::get_legend(plot_end_year(26,1))

plot.count.end.year.26 = plot_end_year(26,1)
plot.count.end.year.28 = plot_end_year(28,1)
plot.count.end.year.30 = plot_end_year(30,1)
```

## Plot total change map 
```{r}
library(MetBrewer)

plot_total_change = function(threshold_chosen,legend_chosen=0){
  p = ggplot() +
    geom_polygon(data=us_aea,aes(x=long,y=lat,group=group),fill='grey35',color='black',size=0.5) +
    geom_polygon(data=USA.df.merged,aes(x=long,y=lat,group=group,color=get(paste0('total_change_wbgt_',threshold_chosen))),size=1) +
    scale_color_gradientn(colors=rev(MetBrewer::met.brewer('Cassatt2')), limits = c(-1*max(USA.df.merged$total_change_wbgt_28),max(USA.df.merged$total_change_wbgt_28))) +
    coord_fixed() + xlab('') + ylab('') + theme_bw() +
    theme(panel.grid.major = element_blank(),
        axis.text.x = element_text(angle=90), axis.ticks.x=element_blank(), legend.text=element_text(size=5),
        panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"),
        panel.border = element_rect(colour = "black"),strip.background = element_blank(),legend.justification='center',
        legend.position = 'bottom', legend.background = element_rect(fill="white", size=.5, linetype="dotted")) +
    theme_map()
  
  if(legend_chosen==0){
    p = p +
          guides(color=FALSE)
  }
  if(legend_chosen==1){
    p = p +
      guides(color = guide_colorbar(direction = "horizontal", title.position="left",barwidth = 10,barheight = 1,title.vjust = 0.8, 
                                  title = legend_create_total_change(threshold_chosen)),legend.text=element_text(size=5))
  }
  
  return(p)
}

plot.total.change.legend = cowplot::get_legend(plot_total_change(26,1))

plot.total.change.26 = plot_total_change(26,1)
plot.total.change.28 = plot_total_change(28,1)
plot.total.change.30 = plot_total_change(30,1)
```

## Plot count in last 5 years map 
```{r}
plot_last_5_years = function(threshold_chosen,legend_chosen=0){
  p = ggplot() +
    geom_polygon(data=us_aea,aes(x=long,y=lat,group=group),fill='grey35',color='black',size=0.5) +
    geom_polygon(data=USA.df.merged,aes(x=long,y=lat,group=group,
                                        color=get(paste0('wbgt_',threshold_chosen,'_',(end_year_wbgt-4),'_',end_year_wbgt))),size=1) +
    scale_color_gradientn(colors=MetBrewer::met.brewer('OKeeffe2'), limits = c(0,max(USA.df.merged$wbgt_28_2016_2020))) +
    coord_fixed() + xlab('') + ylab('') + theme_bw() +
    theme(panel.grid.major = element_blank(),
        axis.text.x = element_text(angle=90), axis.ticks.x=element_blank(),legend.text=element_text(size=5),
        panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"),
        panel.border = element_rect(colour = "black"),strip.background = element_blank(),legend.justification='center',
        legend.position = 'bottom', legend.background = element_rect(fill="white", size=.5, linetype="dotted")) +
    theme_map()
  
  if(legend_chosen==0){
    p = p +
          guides(color=FALSE)
  }
  if(legend_chosen==1){
    p = p +
          guides(color = guide_colorbar(direction = "horizontal", title.position="left",barwidth = 10,barheight = 1,title.vjust = 0.8, 
                                  title = legend_create_count_last_5_years(threshold_chosen)),legend.text=element_text(size=5)) 
    }  

    return(p)

}

plot.count.end.year.legend = cowplot::get_legend(plot_last_5_years(26,1))

plot.count.last.5.years.26 = plot_last_5_years(26,1)
plot.count.last.5.years.28 = plot_last_5_years(28,1)
plot.count.last.5.years.30 = plot_last_5_years(30,1)
```

## Plot state-wide population-weighted annual difference
```{r}
plot_state_year_difference = function(threshold_chosen,legend_chosen=0){
  p = ggplot() +
    geom_tile(data=dat.wbgt.summarised.merged.weighted.prison.state,
                 aes(x=year,y=STATE,fill=get(paste0('wbgt_',threshold_chosen,'_diff'))),size=1) +
    guides(fill = guide_colorbar(direction = "horizontal", title.position="left",barwidth = 10, 
                                 barheight = 1,title.vjust = 0.8,
                                 title = "Difference in number of hot-humid days"),
                                 legend.text=element_text(size=15)) +
    xlab('Year') +
    ylab('State') +
    scale_fill_gradient2(low=met.brewer('Demuth')[10],mid='white',high=met.brewer('Demuth')[1],
                         na.value="white",guide = guide_legend(nrow = 1)) +
    theme_bw() + theme(text = element_text(size = 10),
    panel.grid.major = element_blank(),axis.text.x = element_text(angle=90 , size=8, vjust=0.5),
    plot.title = element_text(hjust = 0.5),panel.background = element_blank(),
    panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"),
    panel.border = element_rect(colour = "black"),strip.background = element_blank(),
    legend.position = 'bottom',legend.justification='center',
    legend.background = element_rect(fill="white", size=.5, linetype="dotted"))
  
  if(legend_chosen==0){
    p = p +
          guides(fill=FALSE)
  }
  if(legend_chosen==1){
    p = p + 
      guides(fill = guide_colorbar(direction = "horizontal", title.position="left",barwidth = 10, 
                                 barheight = 1,title.vjust = 0.8,
                                 title = "Population-weighted difference in annual number of\n hot-humid days between prisons and rest of state"),legend.text=element_text(size=15))
  }
  
  return(p)
}

plot.heatmap.legend = cowplot::get_legend(plot_state_year_difference(28,1))

plot.state.year.diff.26 = plot_state_year_difference(26,1)
plot.state.year.diff.28 = plot_state_year_difference(28,1)
plot.state.year.diff.30 = plot_state_year_difference(30,1)
```

## 26°C: Place arrangement of plots
```{r}
pdf(paste0(wbgtmax.folder,'Figure_2_26.pdf'),paper='a4r',width=0,height=0)
gridExtra::grid.arrange(as_grob(plot.state.year.diff.26),
                        as_grob(plot.total.change.26),
                        widths = c(4, 4),
                        layout_matrix = rbind(c(1, 2),
                                              c(1, 2)),
                        nrow=2) 
dev.off()

gridExtra::grid.arrange(as_grob(plot.state.year.diff.26),
                        as_grob(plot.total.change.26),
                        widths = c(4, 4),
                        layout_matrix = rbind(c(1, 2),
                                              c(1, 2)),
                        nrow=2) 
```

## 28°C: Place arrangement of plots
```{r}
pdf(paste0(wbgtmax.folder,'Figure_2_28.pdf'),paper='a4r',width=0,height=0)
gridExtra::grid.arrange(as_grob(plot.state.year.diff.28),
                        as_grob(plot.total.change.28),
                        widths = c(4, 4),
                        layout_matrix = rbind(c(1, 2),
                                              c(1, 2)),
                        nrow=2) 
dev.off()

gridExtra::grid.arrange(as_grob(plot.state.year.diff.28),
                        as_grob(plot.total.change.28),
                        widths = c(3, 4),
                        layout_matrix = rbind(c(1, 2),
                                              c(1, 2)),
                        nrow=2)
```

## Save output data for Figure 2a
```{r}
dat.wbgt.summarised.merged.weighted.prison.state.output = dat.wbgt.summarised.merged.weighted.prison.state %>%
  dplyr::select(state,year,wbgt_28_diff) %>%
  dplyr::mutate(wbgt_28_diff=round(wbgt_28_diff,1))

readr::write_csv(dat.wbgt.summarised.merged.weighted.prison.state.output ,paste0(wbgtmax.folder,'Figure_2a.csv'))
```

## Save output data for Figure 2b
```{r} 
dat.wbgt.summarised.regression.output = dat.wbgt.summarised.regression %>%
  dplyr::select(prison_id,total_change_wbgt_28) %>%
  dplyr::mutate(total_change_wbgt_28=round(total_change_wbgt_28,1)) %>%
  left_join(shapefile.data,by=c('prison_id'='FID')) %>% 
  filter(STATUS=='OPEN') %>% filter(POPULATION > 0) %>%
  dplyr::select(prison_id,total_change_wbgt_28,NAME,STATE,STATEFP,COUNTY,CITY,TYPE,POPULATION,CAPACITY,VAL_DATE)

readr::write_csv(dat.wbgt.summarised.regression.output ,paste0(wbgtmax.folder,'Figure_2b.csv'))
```