---
title: "EDA facilities"
output: html_document
---

## First step to load packages and file locations etc.
```{r include=FALSE}
rm(list=ls())
project.folder = paste0(print(here::here()),'/')
source(paste0(project.folder,'create_folder_structure.R'))
source(paste0(functions.folder,'script_initiate.R'))
```

## Load specific plotting packages
```{r, message=FALSE}
# load required packages
library(maptools)
library(mapproj)
library(rgeos)
library(rgdal)
library(RColorBrewer)
library(ggplot2)
library(raster)
library(sp)
library(plyr)
library(graticule)
library(zoo)
library(purrr)
library(cowplot)
library(janitor)
library(tidyverse)
```

## Prepare map structure
```{r}
# for map theme to plot in ggplot
theme_map = function(base_size=8 , base_family=""){
    require(grid)
    theme_bw(base_size=base_size,base_family=base_family) %+replace%
    theme(axis.line=element_blank(),
    axis.text=element_blank(),
    axis.ticks=element_blank(),
    axis.title=element_blank(),
    panel.background=element_blank(),
    panel.border=element_blank(),
    panel.grid=element_blank(),
    panel.margin=unit(0,"lines"),
    plot.background=element_blank(),
    legend.position = 'bottom'
    )
}
```

## Load US prison shapefile for mapping
```{r}
# load shapefile of entire United States by prison
us.main = readOGR(dsn=paste0(project.folder,"data/shapefiles/Prison_Boundaries/"),
                           layer="Prison_Boundaries_Edited")

us.main = spTransform(us.main, CRS("+proj=laea +lat_0=45 +lon_0=-100 +x_0=0 +y_0=0 +a=6370997 +b=6370997 +units=m +no_defs"))

# fortify to prepare for plotting in ggplot
map = fortify(us.main)

# extract data from shapefile
us.main@data$id = rownames(us.main@data) 
shapefile.data = us.main@data

# fix the weird Alabama mistake
shapefile.data = shapefile.data %>%
  dplyr::mutate(STATEFP=case_when(STATEFP=='10' & STATE=='AL'~ '01',
                               TRUE ~ STATEFP))

# merge selected data to map_create dataframe for colouring of ggplot
USA.df = merge(map, shapefile.data, by='id')
USA.df$prison_id = as.integer(as.character(USA.df$FID))
```

## Only choose facilities that are operational and with more than zero population
```{r}
shapefile.data = shapefile.data %>% 
  dplyr::filter(STATUS=='OPEN') %>% 
  dplyr::filter(POPULATION > 0) %>%
  dplyr::rename(Type=TYPE) %>%
  dplyr::mutate(Type1=case_when(Type=='COUNTY' ~ 'County',
                               Type=='LOCAL' ~ 'Local',
                               Type=='FEDERAL' ~ 'Federal',
                               Type=='STATE' ~ 'State',
                               TRUE ~ 'Other'))

n_distinct(shapefile.data$FACILITYID)
```

## Join facility and temperature data
```{r}
temperature_data <- readRDS(paste0(wbgt.folder, "weighted_area_raster_prison_wbgtmax_daily_1982_2020_over_time.rds"))

prison_temp_data <- shapefile.data %>% janitor::clean_names() %>% 
  filter(status == "OPEN") %>% 
  dplyr::select(fid, name, address, city, state, type1, population, county) %>% 
  left_join(temperature_data, by = c("fid" = "prison_id"))

prison_pop_state <- prison_temp_data %>% #split this up into two because group_by wasn't working? 
  filter(year == "2018") %>% 
  mutate(total_pop = sum(population),
         pert_pop = population/total_pop) %>% 
  filter(type1 == "State") %>% 
  group_by(state) %>% 
  dplyr::summarize(perc_total_prison_pop = sum(pert_pop),
            total_pop = sum(population))

```

## Number of person-days with dangerous heat
```{r}
fig1a_data <- read_csv(paste0(wbgtmax.folder, 'Figure_1a.csv'))

num_person_days <- fig1a_data %>% 
  filter(Type == "Total") %>% 
  mutate(num_person_days = sum(wbgt_28_pop_mean)) 

num_person_days %>% dplyr::select(num_person_days) %>% distinct()
```

## Percent of total exposure in state facilities
```{r}

state_person_days <- num_person_days %>% dplyr::select(STATE, num_person_days) %>% 
  left_join(fig1a_data) %>% 
  filter(Type == "State") %>% 
  mutate(state_person_days = sum(wbgt_28_pop_mean),
         percent_state_person_days = state_person_days/num_person_days) 
state_person_days %>% dplyr::select(state_person_days, percent_state_person_days) %>% distinct()
```

## Percent of total exposure in county facilities
```{r}
county_person_days <- num_person_days %>% dplyr::select(STATE, num_person_days) %>% 
  left_join(fig1a_data) %>% 
  filter(Type == "County") %>% 
  mutate(county_person_days = sum(wbgt_28_pop_mean),
         percent_county_person_days = county_person_days/num_person_days) 
county_person_days %>% dplyr::select(county_person_days, percent_county_person_days) %>% distinct()
```

## Number of incarcerated people in FL
```{r}
shapefile.data %>% dplyr::filter(STATE == "FL" & Type == "STATE") %>% 
  mutate(incar_state_pop = sum(POPULATION)) %>% 
  dplyr::select(incar_state_pop) %>% distinct()
```

## Number of incarcerated people in TX
```{r}
shapefile.data %>% filter(STATE == "TX" & Type == "STATE") %>% 
  mutate(incar_state_pop = sum(POPULATION)) %>% 
  dplyr::select(incar_state_pop) %>% distinct()
```

## Percent of national incarcerated population in TX and FL state facilities 
```{r}
total_incar_pop_2018 <- sum(shapefile.data$POPULATION)

shapefile.data %>% filter(STATE %in% c("TX", "FL") & Type == "STATE") %>% 
  mutate(incar_state_pop = sum(POPULATION),
         percent_incar = incar_state_pop/total_incar_pop_2018) %>% 
  dplyr::select(percent_incar) %>% distinct()

```

## Percent of national person-days of dangerous humid-hot days in TX and TL state facilities
```{r}
num_person_days %>% 
  filter(STATE %in% c("TX", "FL")) %>% 
  mutate(perc_person_days = wbgt_28_pop_mean/num_person_days,
         sum_perc_person_days = sum(perc_person_days))
```

## Identify facilities with ≥75 hot-humid days, including facility with largest number of h-h days
```{r}
total_days <- prison_temp_data %>% 
  filter(year %in% c(2016,2017,2018,2019,2020)) %>% 
  group_by(fid, name, state, city) %>% 
  dplyr::summarize(total_hot_days = sum(wbgt_28), mean_hot_days = mean(wbgt_28)) %>% 
  filter(mean_hot_days > 74.9999) %>% ungroup()

n_distinct(total_days$fid)

total_days %>% dplyr::slice_max(total_hot_days, n = 1)

```

## Compare incarcerated to non-incarcerated population exposure for US
```{r}
prison_state_comp_data <- read_csv(paste0(wbgtmax.folder,'Figure_2a.csv')) %>% 
  filter(STATE == "USA")
mean(prison_state_comp_data$wbgt_28_diff) 
min(prison_state_comp_data$wbgt_28_diff)
max(prison_state_comp_data$wbgt_28_diff)
```

## Compare incarcerated to non-incarcerated population exposure in AZ and incarcerated in AZ to non-incarcerated national
```{r}
#code from b_02_figure2.Rmd
dat.wbgt.summarised = readRDS(paste0(wbgt.folder,'weighted_area_raster_prison_wbgtmax_daily_',start_year_wbgt,'_',end_year_wbgt,'_over_time.rds'))

dat.wbgt.summarised.merged.weighted.prison = left_join(dat.wbgt.summarised,shapefile.data, by=c('prison_id'='FID')) %>% 
  dplyr::filter(STATUS=='OPEN') %>% 
  dplyr::filter(POPULATION > 0) %>%
  dplyr::group_by(STATE,STATEFP,year) %>%
  dplyr::summarise(wbgt_26 = weighted.mean(wbgt_26,POPULATION),
            wbgt_28 = weighted.mean(wbgt_28,POPULATION),
            wbgt_30 = weighted.mean(wbgt_30,POPULATION),
            wbgt_35 = weighted.mean(wbgt_35,POPULATION)) 

STATES = dat.wbgt.summarised.merged.weighted.prison %>% pull(STATE) %>% unique()

dat.wbgt.summarised.merged.weighted.prison.national = left_join(dat.wbgt.summarised,shapefile.data, by=c('prison_id'='FID')) %>% 
  dplyr::filter(STATUS=='OPEN') %>% 
  dplyr::filter(POPULATION > 0) %>%
  dplyr::group_by(year) %>%
  dplyr::summarise(wbgt_26 = weighted.mean(wbgt_26,POPULATION),
            wbgt_28 = weighted.mean(wbgt_28,POPULATION),
            wbgt_30 = weighted.mean(wbgt_30,POPULATION),
            wbgt_35 = weighted.mean(wbgt_35,POPULATION)) %>%
  dplyr::mutate(STATE='USA',STATEFP='99') %>%
  dplyr::relocate(STATE) %>%
  dplyr::relocate(STATEFP)

dat.wbgt.summarised.merged.weighted.prison = 
  rbind(dat.wbgt.summarised.merged.weighted.prison,dat.wbgt.summarised.merged.weighted.prison.national) %>%
  mutate(STATE=factor(STATE, levels=rev(c('USA',STATES))))

dat.wbgt.summarised.merged.weighted.state = readRDS(paste0(wbgt.folder,'weighted_area_raster_state_wbgtmax_daily_',start_year_wbgt,'_',end_year_wbgt,'_over_time.rds'))

dat.wbgt.summarised.merged.weighted.prison = dat.wbgt.summarised.merged.weighted.prison %>%
  dplyr::rename(wbgt_26_prison=wbgt_26,wbgt_28_prison=wbgt_28,wbgt_30_prison=wbgt_30,wbgt_35_prison=wbgt_35) %>%
  dplyr::rename(state=STATEFP)

dat.wbgt.summarised.merged.weighted.state = dat.wbgt.summarised.merged.weighted.state %>%
  dplyr::rename(wbgt_26_state=wbgt_26,wbgt_28_state=wbgt_28,wbgt_30_state=wbgt_30,wbgt_35_state=wbgt_35)

dat.wbgt.summarised.merged.weighted.prison.state = left_join(dat.wbgt.summarised.merged.weighted.prison,
                                                             dat.wbgt.summarised.merged.weighted.state,
                                                             by=c('state','year')) %>%
  dplyr::mutate(wbgt_26_diff = wbgt_26_prison - wbgt_26_state,
                wbgt_28_diff = wbgt_28_prison - wbgt_28_state,
                wbgt_30_diff = wbgt_30_prison - wbgt_30_state,
                wbgt_35_diff = wbgt_35_prison - wbgt_35_state) %>% 
  filter(STATE == "AZ")

mean(dat.wbgt.summarised.merged.weighted.prison.state$wbgt_28_diff)

mean_az_prison <- mean(dat.wbgt.summarised.merged.weighted.prison.state$wbgt_28_prison)

#load non-prison (state) data
non_prison_data <- readRDS(paste0(wbgt.folder, "weighted_area_raster_state_wbgtmax_daily_1982_2020_over_time.rds")) 
non_prison_spf <- non_prison_data %>% filter(state == "99")
mean_natl_non_prison <- mean(non_prison_spf$wbgt_28)

mean_az_prison - mean_natl_non_prison 
```

## Number of people in facilities with trends in hot-humid days
```{r}
prison_trend_data <- read_csv(paste0(wbgtmax.folder,'Figure_2b.csv')) %>% janitor::clean_names() %>% 
  mutate(total_incar_pop = sum(population),
         pert_total_incar = population/total_incar_pop)

prison_trend_data %>% filter(total_change_wbgt_28 > 0) %>% tally(population)
prison_trend_data %>% filter(total_change_wbgt_28 > 0) %>% tally(pert_total_incar)
prison_trend_data %>% filter(total_change_wbgt_28 > 0) %>% count()
```

## Number of people in FL facilites with trends in hot-humid days
```{r}
specific_regions <- prison_trend_data %>% 
  filter(state == "FL")

mean(specific_regions$total_change_wbgt_28)
```
```{r}
prison_diff_data <- read_csv(paste0(wbgtmax.folder,'Figure_2c.csv')) %>% janitor::clean_names() 

prison_diff_data %>% dplyr::slice_max(total_change_disparity_wbgt_28, n = 1)

```

